---
title: "Get started with the igoR package"
author: "Diego Hernangómez"
bibliography: refs_igo.bib
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with the igoR package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitropts, include=FALSE}

start_time <- Sys.time()

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.width = 6,
  fig.height = 4,
  out.width = "100%",
  out.height = "60%"
)
```

This vignette is meant to provide **useRs** with an visual, explorable
introduction to the capabilities of the `igoR` package.

The analysis would be based on those provided on
@igo_2020. For more information on the IGO data sets and
additional downloads, see [Intergovernmental Organizations
(v3)](https://correlatesofwar.org/data-sets/IGOs).

*Note that the dyadic dataset is not provided in the package, due its size
(\~500 MB on Stata `.dta` format). However,* `igo_dyadic()` *function provides
similar results.*

## Definitions

From @codebook_igo_2019:

### What is an IGO?

The definition of an Intergovernmental Organization (IGO) on the original
dataset is based on the following criteria:

1.  An IGO must consist of at least three members of the [COW-defined state
    system](https://correlatesofwar.org/data-sets/cow-country-codes).

2.  An IGO must hold regular plenary sessions at least once every ten years

3.  An IGO must possess a permanent secretariat and corresponding headquarters.

### When does an IGO actually begin?

The data sets begins to code an IGO by identifying the first year in which the
organization functions. In some cases, individual members are listed by year of
accession or signature.

### When does an IGO die?

Version 3.0 of the IGO data set uses the following criteria:

-   An organization is considered terminated when the following words were used
    to describe the context of the organization:

    -   Replaced;
    -   Succeeded;
    -   Superseded;
    -   Integrated;
    -   Merged;
    -   Dies.

## Analysis

This section provides some quick analysis based on the figures of 
@igo_2020.

### Initial Setup

```{r setup}

library(igoR)

# Additional libraries
library(ggplot2)
library(dplyr)
```

<details><summary><strong>Custom <code>ggplot2</code> theme</strong>: <code>theme_igoR</code></summary>

```{r custom_theme}

theme_igoR <- theme(
  axis.title = element_blank(),
  axis.line.x.bottom = element_line("black"),
  axis.line.y.left = element_line("black"),
  axis.text = element_text(color = "black", family = "sans"),
  axis.text.y.left = element_text(angle = 90),
  legend.position = "bottom",
  legend.title = element_blank(),
  legend.key = element_blank(),
  legend.key.width = unit(2, "cm"),
  legend.text = element_text(family = "sans", size = 13),
  legend.box.background = element_rect(color = "black", size = 1),
  legend.spacing = unit(1.2 / 100, "npc"),
  plot.background = element_rect("grey90"),
  plot.margin = unit(rep(0.5, 4), "cm"),
  panel.background = element_rect("white"),
  panel.grid = element_blank(),
  panel.border = element_rect(fill = NA, colour = "grey90"),
  panel.grid.major.y = element_line("grey90")
)
```

</details>

### IGOs overview

The following code extracts the number of IGOs and states included on this
package. The years available are 1816 to 2014.

```{r Fig1, message=FALSE, warning=FALSE, fig.cap="Figure 1. IGOs and states in the world system, 1816-2014"}

# Summarize
igos_by_year <- igo_year_format3 %>%
  group_by(year) %>%
  summarise(value = n(), .groups = "keep") %>%
  mutate(variable = "Total IGOs")

countries_by_year <- state_year_format3 %>%
  group_by(year) %>%
  summarise(value = n(), .groups = "keep") %>%
  mutate(variable = "Number of COW states")

all_by_year <- rbind(igos_by_year, countries_by_year) %>%
  mutate(variable = factor(variable))

# Reverse values
all_by_year$variable <- factor(all_by_year$variable,
  levels = rev(levels(all_by_year$variable))
)

# Plot
ggplot(all_by_year, aes(x = year, y = value)) +
  geom_line(size = 0.8, color = "black", aes(linetype = variable)) +
  scale_linetype_manual(values = c("solid", "dashed")) +
  geom_vline(xintercept = c(1945, 1989)) +
  ylim(0, 400) +
  theme_igoR
```

### IGO Births and deaths

This plot shows how many IGOs were "born" and "died" on each year

```{r Fig2, message=FALSE, warning=FALSE, fig.cap="Figure 2. Birth and death rates of IGOs, 1816-2014"}

# Births and deads by year

df <- igo_search()

births <- df %>%
  mutate(year = sdate) %>%
  group_by(year) %>%
  summarise(value = n(), .groups = "keep") %>%
  mutate(variable = "IGO Births")

deads <- df %>%
  mutate(year = deaddate) %>%
  group_by(year) %>%
  summarise(value = n(), .groups = "keep") %>%
  mutate(variable = "IGO Deaths")


births_and_deads <- rbind(births, deads) %>% filter(!is.na(year))


# Plot
ggplot(births_and_deads, aes(x = year, y = value)) +
  geom_line(size = 0.8, color = "black", aes(linetype = variable)) +
  scale_linetype_manual(values = c("solid", "dashed")) +
  ylim(0, 15) +
  theme_igoR
```

### IGOs across regions

A plot with the number of IGOs by region. The definition of region is based on
the original definition by @igo_2020.

<details><summary>
<strong>Regions definition</strong>
</summary>

```{r regions_code}
# Extracted from analysis-jpr.do - See https://www.prio.org/JPR/Datasets/

# crossreg and universal codes not included

Asia <- c(
  550, 560, 570, 580, 590, 600, 610, 640, 650, 660,
  670, 725, 750, 825, 1030, 1345, 1400, 1530, 1532, 2300,
  2770, 3185, 3330, 3560, 3930, 4115, 4150, 4160, 4170,
  4190, 4200, 4220, 4265, 4440
)

MiddleEast <- c(
  370, 380, 390, 400, 410, 420, 430, 440, 450, 460,
  470, 490, 500, 510, 520, 1110, 1410, 1990, 2000,
  2220, 3450, 3800, 4140, 4270, 4380
)

Europe <- c(
  20, 300, 780, 800, 832, 840, 860, 1020, 1050, 1070, 1080,
  1125, 1140, 1390, 1420, 1440, 1563, 1565, 1580, 1585, 1590,
  1600, 1610, 1620, 1630, 1640, 1645, 1653, 1660, 1670, 1675,
  1680, 1690, 1700, 1710, 1715, 1720, 1730, 1740, 1750, 1760,
  1770, 1780, 1790, 1800, 1810, 1820, 1830, 1930, 1970, 1980,
  2310, 2325, 2345, 2440, 2450, 2550, 2575, 2610, 2650, 2705,
  2890, 2972, 3010, 3095, 3230, 3290, 3360, 3485, 3505, 3585,
  3590, 3600, 3610, 3620, 3630, 3640, 3650, 3655, 3660, 3665,
  3762, 3810, 3855, 3860, 3910, 4000, 4350, 4450, 4460, 4510,
  4520, 4540
)

Africa <- c(
  30, 40, 50, 60, 80, 90, 100, 110, 115, 120, 125, 130, 140,
  150, 155, 160, 170, 180, 190, 200, 210, 225, 240, 250, 260, 280,
  290, 690, 700, 710, 940, 1060, 1150, 1170, 1260, 1290, 1310,
  1320, 1330, 1340, 1355, 1430, 1450, 1460, 1470, 1475, 1480,
  1500, 1510, 1520, 1870, 2080, 2090, 2230, 2330, 2795, 3300,
  3310, 3470, 3480, 3510, 3520, 3570, 3740, 3760, 3761, 3790,
  3820, 3875, 3905, 3970, 4010, 4030, 4050, 4055, 4080, 4110,
  4120, 4130, 4230, 4240, 4250, 4251, 4340, 4365, 4480, 4485,
  4490, 4500, 4501, 4503
)

Americas <- c(
  310, 320, 330, 340, 720, 760, 815, 875, 880, 890, 900,
  910, 912, 913, 920, 950, 970, 980, 990, 1000, 1010, 1095,
  1130, 1486, 1489, 1490, 1860, 1890, 1920, 1950, 2070, 2110,
  2120, 2130, 2140, 2150, 2160, 2170, 2175, 2180, 2190, 2200,
  2203, 2206, 2210, 2260, 2340, 2490, 2560, 2980, 3060,
  3340, 3370, 3380, 3390, 3400, 3410, 3420, 3428, 3430, 3670,
  3680, 3812, 3830, 3880, 3890, 3900, 3925, 3980, 4070, 4100,
  4260, 4280, 4370
)

regions <- igo_search()

regions$region <- NA

regions <- regions %>% select(region, ionum)

regions$region <-
  ifelse(regions$ionum %in% Africa, "Africa", regions$region)

regions$region <-
  ifelse(regions$ionum %in% Americas, "Americas", regions$region)

regions$region <-
  ifelse(regions$ionum %in% Asia, "Asia", regions$region)

regions$region <-
  ifelse(regions$ionum %in% Europe, "Europe", regions$region)

regions$region <-
  ifelse(regions$ionum %in% MiddleEast, "Middle East", regions$region)
```

</details>

After we have created a data frame with the regions, we can classify the IGOs by
region.

```{r Fig3, message=FALSE, warning=FALSE, fig.cap="Figure 3. IGO counts across regions, 1816-2014"}

# regions dataset created on previous chunk

# All IGOs
alligos <- igo_year_format3[, c("year", "ionum")]

regionsum <- merge(alligos, regions) %>%
  group_by(year, region) %>%
  summarise(value = n(), .groups = "keep") %>%
  filter(!is.na(region))

# Order
regionsum$region <- factor(regionsum$region,
  levels = c(
    "Asia", "Europe",
    "Africa", "Americas",
    "Middle East"
  )
)

# Plot
ggplot(regionsum, aes(x = year, y = value)) +
  geom_line(size = 0.8, color = "black", aes(linetype = region)) +
  scale_linetype_manual(values = c(
    "solid", "dashed",
    "dotted", "dotdash", "longdash"
  )) +
  guides(linetype = guide_legend(ncol = 2, byrow = TRUE)) +
  ylim(0, 80) +
  theme_igoR
```

### Selected Countries: Asia

Number of memberships of a country. We select here five countries on Asia:
India, China, Pakistan, Indonesia and Bangladesh.

```{r Fig4, message=FALSE, warning=FALSE, fig.cap="Figure 4. IGO membership: five states in Asia, 1865-2014"}

asia5_cntries <- c(
  "China", "India",
  "Pakistan", "Indonesia",
  "Bangladesh"
)

# Five countries of Asia
asia5_igos <- igo_state_membership(
  state = asia5_cntries,
  year = 1865:2014,
  status = c("Full Membership")
)

asia5 <- asia5_igos %>%
  group_by(statenme, year) %>%
  summarise(values = n(), .groups = "keep")


# Reorder
asia5$statenme <- factor(asia5$statenme,
  levels = asia5_cntries
)



# Plot
ggplot(asia5, aes(x = year, y = values)) +
  geom_line(size = 0.8, color = "black", aes(linetype = statenme)) +
  scale_linetype_manual(values = c(
    "solid", "dashed",
    "dotted", "dotdash",
    "longdash"
  )) +
  guides(linetype = guide_legend(ncol = 3, byrow = TRUE)) +
  theme(axis.title.y.left = element_text(
    family = "sans",
    size = 12,
    margin = margin(r = 6)
  )) +
  scale_y_continuous(
    "Number of memberships",
    breaks = c(0, 20, 40, 60, 80, 100),
    limits = c(0, 95),
    labels = as.character(c(0, 20, 40, 60, 80, 100))
  ) +
  theme_igoR
```

### Shared memberships

Number of shared full memberships between Spain and four selected countries:

```{r Fig5, message=FALSE, warning=FALSE, fig.cap="Figure 5. Number of IGOs with full shared memberships with Spain (selected countries), 1816-2014"}

selected_countries <- c(
  "France",
  "Morocco",
  "China",
  "USA"
)

Spain_Selected <- igo_dyadic("Spain", selected_countries)

# Compute number of shared memberships

Spain_Selected$values <- rowSums(Spain_Selected == 1)

# Plot
ggplot(Spain_Selected, aes(x = year, y = values)) +
  geom_line(size = 0.8, color = "black", aes(linetype = statenme2)) +
  scale_linetype_manual(values = c(
    "solid", "dashed",
    "dotted", "dotdash"
  )) +
  guides(linetype = guide_legend(ncol = 2, byrow = TRUE)) +
  theme(axis.title.y.left = element_text(
    family = "sans",
    size = 10,
    margin = margin(r = 6)
  )) +
  scale_y_continuous(
    "Shared memberships w\ Spain",
    breaks = c(0, 20, 40, 60, 80, 100),
    limits = c(0, 110),
    labels = as.character(c(0, 20, 40, 60, 80, 100))
  ) +
  theme_igoR +
  geom_vline(xintercept = 1939, alpha = 0.2) +
  annotate(
    "label",
    x = 1938,
    y = 60,
    size = 3,
    label = "Spanish \nCivil War"
  ) +
  geom_vline(xintercept = 1978, alpha = 0.2) +
  annotate(
    "label",
    x = 1970,
    y = 100,
    size = 3,
    label = "Constitution\nof Spain"
  )
```

## References
