---
title: "Mapping IGOs"
bibliography: refs_igo.bib
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mapping IGOs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# Install if not already

if (isFALSE(requireNamespace("giscoR", quietly = TRUE))) {
  install.packages("giscoR")
}

```

This vignette provides some geospatial visualizations using the IGO datasets (@doi:10.1177/0022343319881175) included in this package.

```{r setup, warning=FALSE}

library(igoR)
library(dplyr)
library(countrycode)

# Geospatial packages
library(sf)
library(giscoR)
library(ggplot2)


```


## Evolution of the composition of UN

A map showing the evolution of countries that are members of the United
Nations:

```{r UN}
un <- igo_search("UN", exact = TRUE)

knitr::kable(un)

# Extract three dates
UN1950 <- igo_members("UN", 1950)
UN1980 <- igo_members("UN", 1980)
UN2010 <- igo_members("UN", 2010)

# Create a single dataframe

UN <-
  UN1950 %>% rbind(UN1980) %>% rbind(UN2010) %>%
  select(ccode, state, year) %>% group_by(ccode, state) %>%
  summarise(year = min(year), .groups = "keep")

UN$year <- factor(UN$year)
# Extract ISO3Code
UN$ISO3_CODE <- countrycode(UN$ccode, "cown", "iso3c")

knitr::kable(UN %>% filter(is.na(ISO3_CODE)))


# World in Robinson projection
world <-
  gisco_get_countries(year = 2010) %>% st_transform("+proj=robin +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs")

UNmap <- merge(world, UN)

ggplot(world) + geom_sf(fill = "grey80", color = NA) +
  geom_sf(data = UNmap, color = NA, aes(fill = year)) +
  scale_fill_manual(values = adjustcolor(c("#3690C0", "#74A9CF", "#023858"),alpha.f = 0.8)) +
  ggtitle("UN members") +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 15),
    legend.position = "bottom",
    panel.grid = element_line(colour = "grey80"),
    legend.title = element_blank(),
    legend.key = element_blank(),
  )



```

# Members of the OAPEC

A map showing the members of the Organization of Arab Petroleum Exporting Countries (OAPEC):

```{r OAPEC}

OAPEC <- igo_members("OAPEC")
OAPEC$ISO3_CODE <- countrycode(OAPEC$ccode, "cown", "iso3c")

# World
world2 <- gisco_get_countries(epsg = 3035)

OAPECmap <- merge(world2, OAPEC)

limits <- st_bbox(OAPECmap)

ggplot(world2) + geom_sf(fill = "#B9B9B9", color = NA) +
  geom_sf(data = OAPECmap, color = "#335033", aes(fill = ioname)) +
  scale_fill_manual(values = "#346733") +
  coord_sf(xlim = limits[c(1, 3)],
           ylim = limits[c(2, 4)]) + theme_void() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.key = element_blank(),
  )

```

## Session Info

<details>
  <summary><strong>Details</strong></summary>
```{r session_info, echo=FALSE}
sessionInfo()

```
</details>



## References
